---
title: "Homework 6: Functional Programming"
output: html_notebook
author: "Kamryn Parker"
---
```{r}
library(tidyverse)
library(palmerpenguins)
```

1. Using map_dbl, compute the mean of every numeric column in diamonds (from the tidyverse library).
```{r}
diamonds %>%
   select_if(is.numeric) %>% #only choosing the numeric columns
   map_dbl(mean, na.rm=TRUE)
```

2. Using map_chr, determine the type (using the typeof() function) of each column in diamonds.
```{r}
map_chr(diamonds, typeof) #Shows what each type of column is
```

3. Using map_int, count the number of unique values in each column of penguins. (Hint: There's an example of this in my lecture code from Thursday)
```{r}
penguins %>%
  map_int(function(x) length(unique(x))) #create an anonymous function x and map it to unique values in penguins and use length to count them
```

4a. Define a function that takes two arguments, x and y, adds them together, and returns a text string that reads "x + y = result". In other words, your_function(1,2) will return the pasted string: "1 + 2 = 3".
```{r}
new_function <- function(x,y){ #takes in values x and y
  z <- x + y #use variable z as the "result" of function variables
  (paste(x, "+", y, '=',z)) #using paste function to fit the string conditions
}
new_function(1,2) #calls the functions

```
4b. Using map2_chr, apply your new function to the depth and table columns from diamonds. Save the results into a new column. You'll need to use the mutate function for this (see the example in the code from Thursday).
```{r}
diamonds %>%
  mutate(new_function_column = map2_chr(depth, table, new_function)) %>% #Creates a new column in entire dataframe
  select(new_function_column) #selects that new column
```

5. Nest diamonds by cut
```{r}
diamonds %>%
  group_by(cut) %>% #groups the dataframe by cut
  nest() %>% #nests the rest of the data frame inside of the cut variables
  View()
```
6. Split diamonds by cut
```{r}
diamonds %>%
  split(.$cut) #splits up each cut variable and spits it out into its own dataframe
```

